---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# capstoner

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/hugostar7/capstoner.svg?branch=main)](https://travis-ci.com/hugostar7/capstoner)
[![R-CMD-check](https://github.com/hugostar7/capstoner/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hugostar7/capstoner/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of capstoner is to provide means to analyse earthquakes data. Functions in this package can be used to create data visualization tools time series plots and maps of earthquakes data with specific columns. 
The functions eq_clean_data() and eq_location_clean() can be used to clean  the data and create the columns necessary(DATE, COUNTRY,
LOCATION_NAME) for our data analysis. The function geom_timeline() creates time series plots which can be  annotated with geom_timeline_label(). The function eq_create_label() helps create popups on interactive maps  created using eq_maps().

## Installation

You can install the development version of capstoner from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hugostar7/capstoner")
```

## Example

This is a basic example which shows you how to clean your data using
`eq_clean_data` and `eq_location_clean`

```{r data_cleaning}
library(capstoner)
library(magrittr)

data("earthquakes")

# Clean data and create DATE column
cd = eq_clean_data(earthquakes)
head(cd)[ , 33:37]

# Clean data and create proper COUNTRY and LOCATION_NAME columns
lcd = eq_location_clean(earthquakes)
head(lcd)[ , 35:38]
```

You can also create an interactive map for the earthquakes data:

```{r interactiveMap}
earthquakes %>%
  eq_location_clean() %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  eq_map(annot_col = "DATE")

# You can also use eq_location_clean to filter the data
eq_location_clean(earthquakes, year_min = 2000, countries = "MEXICO") %>%
  eq_map(annot_col = "DATE")
```

Let's  customize the popups on the previous map using
`eq_create_label`

```{r custm_popup}
earthquakes %>%
  eq_location_clean() %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```

